State of the Practice for Lattice Boltzmann Method Software

### 2022-04-21 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits to introduction section

A small sample of important applications of LBM includes the following:
designing fuel cells \citep{ZhangEtAl2018}, modelling groundwater flow
\citep{AnwarAndSukop2009}, and analyzing the flow of blood in the cardiovascular
system \citep{SadeghiEtAl2022b, SadeghiEtAl2022, SadeghiEtAl2020}.  As these
examples illustrate, LBM can be used for tackling problems that impact such
areas as environmental policy, manufacturing, and health and safety.  Given the
important applications of LBM, users of the libraries will be concerned with
software qualities like reliability, robustness, reproducibility, performance,
correctness and verifiability.  Since their time is valuable, developers that
create or modify LBM libraries will have additional quality concerns, including
maintainability, reusability, and understandability.  With the quality
considerations in mind, and the potentially overwhelming number of choices for
existing LBM libraries, the time is right to assess the state of the practice.
The goal of this report is to analyze the current state of LBM software
development and compare it to the development practices employed for research
software in general.  We want to highlight success stories that LBM developers
can share amongst their community, and with the broader research software
community, while at the same time looking for areas for potential future
improvement.

To focus our efforts, we devised 10 research questions.  The questions are used
to structure the discussion in the paper, so for each research question below we
point to the section that contains our answer.  The questions are inspired by
the research questions that underpin our data collection methodology
\citep{SmithEtAl2021, SmithAndMichalski2022}. We start with identifying the
examples of LBM software where we have access to the source code:

@@

We next wish to assess the representative software to determine how well they
apply current software development best practices.  By \emph{best practices} we
mean methods, techniques and tools that are generally believed to improve
software development, like testing and documentation.  As we will discuss in
Section~\ref{empiricalmeasures}, we will categorize our best practices around
software qualities.  At this point in the process, to remove potential
user/developer bias, we will base our assessment only on publicly available
artifacts, where artifacts are the documents, scripts and code that we find in a
project's public repository. Example artifacts include requirements,
specifications, user manuals, unit test cases, system tests, usability tests,
build scripts, API (Application Programming Interface) documentation, READMEs,
license documents, process documents, and code. Following best practices does
not guarantee popularity, so we will also compare our ranking to how the user
community itself ranks the identified projects.

@@

To understand the state of the practice we wish to learn the frequency with
which different artifacts appear, the types of development tools used and the
methodologies used for software development.  With this data, we can ask
questions about how LBM software compares to other research software.  For cases
where data is available, we will compare to other specific domains of research
software, like medical imaging \citep{Dong2021} and ocean modelling
\citep{JungEtAl2022}.  However, most of the literature on the state of the
practice does not focus on specific domains, but rather on all research software
\citep{SmithAndMichalski2022}. We will refer to studies and guidelines that do
not distinguish an application domain as \emph{research software in general}. By
comparing LBM software to research software in general we can determine the ways
LBM follows the trends from other developer communities, and ways in which it is
different.

@@

Only so much information can be gleaned by looking at software repositories.  To
gain additional insight, we need to interview developers.  We need to learn
their concerns, how they deal with these concerns; we need to learn what pain
points exist for them.  We wish to know what practices are used by the top LBM
projects, so that others can potentially emulate these practices. We also wish
to identify new practices that LBM developers can adopt to improve their
software in the future. The above points are covered by the questions outlined
below:

>>>>

A small sample of safety and mission-critical applications of LBM includes the
following: designing fuel cells \citep{ZhangEtAl2018}, modelling groundwater
flow \citep{AnwarAndSukop2009}, and analyzing the flow of blood in the
cardiovascular system \citep{SadeghiEtAl2022b, SadeghiEtAl2022,
SadeghiEtAl2020}.  As these examples illustrate, engineers and scientists use
LBM for tackling problems that impact such areas as environmental policy,
manufacturing, and health and safety. Given the critical applications of LBM,
users of the libraries are concerned with software qualities like reliability,
robustness, reproducibility, performance, correctness and verifiability.  Since
their time is valuable, developers of LBM libraries have additional quality
concerns, including maintainability, reusability, and understandability.  With
the quality considerations in mind, and the potentially overwhelming number of
choices for existing LBM libraries, the time is right to assess the state of the
practice. Therefore, we analyze the current state of LBM
software development and compare it to the development practices employed for
research software in general.  We want to highlight success stories that LBM
developers can share amongst their community, and with the broader research
software community, while at the same time looking for areas for potential
future improvement.

We use ten research questions to structure and focus our discussion. Our data
collection methodology \citep{SmithEtAl2021, SmithAndMichalski2022} inspires the
questions.  For each question below, we point to the section that contains our
answer.  We start with identifying the examples of LBM software where we have
access to the source code:

@@

We next wish to assess the representative software to determine how well they
apply current software development best practices.  By \emph{best practices}, we
mean methods, techniques and tools that are generally believed to improve
software development, like testing and documentation.  We will categorize our
best practices around software qualities, as we will discuss in
Section~\ref{empiricalmeasures}.  At this point in the process, to remove
potential user/developer bias, we will base our assessment only on publicly
available artifacts, where artifacts are the documents, scripts and code that we
find in a project's public repository. Example artifacts include requirements,
specifications, user manuals, unit test cases, system tests, usability tests,
build scripts, API (Application Programming Interface) documentation, READMEs,
license documents, process documents, and code. Following best practices does
not guarantee popularity, so we will also compare our ranking to how the user
community ranks the identified projects.

@@

To understand the state of the practice, we wish to learn the frequency with
which different artifacts appear, the types of development tools used and the
methodologies used for software development.  With this data, we can ask
questions about how LBM software compares to other research software.  For cases
where data is available, we can compare to other specific domains of research
software, like medical imaging \citep{Dong2021} and ocean modelling
\citep{JungEtAl2022}.  However, most of the literature on the state of the
practice does not focus on specific domains, but rather on all research software
\citep{SmithAndMichalski2022}. We will refer to studies and guidelines that do
not distinguish an application domain as \emph{research software in general}. By
comparing LBM software to research software in general we can determine how LBM
follows the trends from other developer communities and how it is different.

@@

Only so much information can be gleaned by looking at software repositories.  To
gain additional insight, we need to interview developers.  We need to learn
their concerns and how they deal with them; we need to understand their pain
points. We wish to know what practices are used by the top LBM projects, so that
others can potentially emulate these practices. We also wish to identify new
practices that LBM developers can adopt to improve their software in the future.
The question below cover these points:
